---
title: "WCVI LRPs"
author: "Carrie Holt"
date: "January 2, 2021"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))  


library(tidyverse)
library(ggplot2)
library(gsl)
library(TMB)
library(viridis)

```

This report contains preliminary LRPs for WCVI Chinook. Only indicators stocks with average PNI > 0.5 were included in these preliminary analyses. Results and LRPs with all indicator stocks (enhanced and unenhanced) are included at the end.  Enhanced indicator stocks are: Burman,  Conuma, Leiner, Nitinat, Sarita,  Somass,  Zeballos,  and San Juan. Non-enhanced inidcators stocks are: Artlish, Bedwell/Ursus, Cayeghle, Gold, Kaouk, Marble, Megin, Moyeha, Nahmint, Tahsis, Tahsish, and Tranquil.


```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 1**. Time series of spawner abundances by inlet for indicator WCVI Chinook stocks with average PNI >0.5. Three reference points are included: SREP from the watershed-area model (horizontal green line) with 95% prediction intervals (green shaded area), and SMSY (horizontal yellow line) and SGEN (horizontal blue line) from watershed-area SREP values and expert-derived estimates of productivity."}
source("helperFunctions.r")
source("WCVILRPs.r")
source("PlotSR.r")

xx <- Get.LRP(remove.EnhStocks=TRUE)

plotWCVI_timeseries(WCVIEsc=xx$WCVIEsc, remove.EnhStocks = TRUE)


```

For each year, CU-level status is red when any one inlet within the CU is below SGEN (lower benchmark) in that year.  If all inlets within a CU are above their lower benchmark, then status is amber or green.

```{r , echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 2**. Annual CU Status, where 0 is red status and 1 is amber or green status"}

par(mfrow=c(3,1), cex=1.1,  mar = c(2, 5, 2, 1) + 0.1)
Years <- rownames(xx$WCVIEsc)

for (i in c(2,1,3)){
  plot(x=Years, y=xx$CU_Status[,i], pch=20, xlab="", ylab="", ylim=c(0,1), las=1 , yaxp=c(0,1,1))
  mtext(colnames(xx$CU_Status)[i], side=3, line=0.5, at="1952", adj=0, cex=1.2)
  if(i==1) mtext("CU Status = Red(0) or Amber/Green(1)", side=2, line=2, cex=1.4)

}

```

SMU-level status is identified from the proportion of CUs that are in the red zone. Only years when status of all three CUs are available are plotted and used in the logistic regression to derive an LRP (n=13).
```{r , echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 3**. Time-series of the proportion of CUs within the SMU that are in the red zone"}


par(mfrow=c(1,1), cex=1.1,  mar = c(2, 3, 2, 1) + 0.1)
plot(x=Years, y=xx$SMU_ppn, pch=20, xlab="", ylab="", ylim=c(0,1), las=1)
mtext("WCVI SMU", side=3, line=0.5, at="1952", adj=0, cex=1.4)

```

The LRP is derived from a logistic regression of aggregate abundances to the SMU (summation of indicator stocks with average PNI > 0.5)  versus the proportion of CUs above the red zone. A penalty is added to the estimation procedure to ensure the y-intercept of the curve is near zero. A provisional LRP at 95% of CUs above their lower benchmark is plotted, but alternatives can be considered (e.g., 2/3 of CUs). This logistic regression has a relatively poor fit to the data; in one year aggregate abundances are > LRP when only 1/3 CUs > red zone, and in 2 years aggregate abundances are < LRP but all CUs are > red zone. Diagnostic statistics and plots are in development.

```{r, echo=FALSE, fig.cap ="**Figure 4**. Logistic regression of aggregate abundances vs proportion of CUs above the red zone, with an LRP at 95% of CUs above red zone (vertical dashed line, +/- 95% CIs = vertical dotted lines). Grey band are the confidence intervals for the fitted proportions of CUs."}

plotLogistic(Data=xx$out$Logistic_Data, Preds=xx$out$Preds, LRP=xx$out$LRP, useGenMean = F, plotName="WCVI_logReg_noEnh", outDir="DataOut", p=0.95)

```

Aggregate SMU abundances vary around the LRP, but are mostly below it. Note, the 95% confidence interval for LRP do not yet include uncertainty from the watershed-area regression. Full uncertainties that capture variabilty from benchamrks estimation will be larger.
```{r , echo=FALSE, fig.cap = "**Figure 5**. Time-series of SMU abundances (summation of indicators stocks with average NI >0.5) relative to LRP."}

plotWCVI_SMUtimeseries(SMU_Esc=xx$SMU_Esc, out=xx$out, WCVI_Esc=xx$WCVIEsc)

```
## Plots with enhancement

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 6**. Time series of spawner abundances by inlet including all indicator WCVI Chinook stocks. Three reference points are included: SREP from the watershed-area model (horizontal green line) with 95% prediction intervals (green shaded area), and SMSY (horizontal yellow line) and SGEN (horizontal blue line) estimated from SREP and expert derived productivity. The reference points in this plot differ from those in Fig. 1 because all indicator stocks were included in the watershed-area model instead of only un-enhanced indicator stocks."}
yy <- Get.LRP(remove.EnhStocks=FALSE)

plotWCVI_timeseries(WCVIEsc=yy$WCVIEsc, remove.EnhStocks = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 7**. Annual CU Status with enhancement, where 0 is red status and 1 is amber or green status"}

par(mfrow=c(3,1), cex=1.1,  mar = c(2, 5, 2, 1) + 0.1)
Years <- rownames(yy$WCVIEsc)

for (i in c(2,1,3)){
  plot(x=Years, y=yy$CU_Status[,i], pch=20, xlab="", ylab="", ylim=c(0,1), las=1 , yaxp=c(0,1,1))
  mtext(colnames(yy$CU_Status)[i], side=3, line=0.5, at="1952", adj=0, cex=1.2)
  if(i==1) mtext("CU Status = Red(0) or Amber/Green(1)", side=2, line=2, cex=1.4)

}

```
```{r , echo=FALSE, fig.cap ="**Figure 8**. Time-series of proportion of CUs within the SMU in the red zone, with enhancement"}


par(mfrow=c(1,1), cex=1.1,  mar = c(2, 3, 2, 1) + 0.1)
plot(x=Years, y=yy$SMU_ppn, pch=20, xlab="", ylab="", ylim=c(0,1), las=1)
mtext("WCVI SMU", side=3, line=0.5, at="1952", adj=0, cex=1.4)


```

```{r , echo=FALSE, fig.cap ="**Figure 9**. Logistic regression of aggregate abundances vs proportion of CUs above the red zone, with an LRP at 95% of CUs above red zone (vertical dashed line, +/- 95% CIs = vertical dotted lines), including enhanced stocks. Grey band are the confidence intervals for the fitted proportions of CUs."}

plotLogistic(Data=yy$out$Logistic_Data, Preds=yy$out$Preds, LRP=yy$out$LRP, useGenMean = F, plotName="WCVI_logReg_noEnh", outDir="DataOut", p=0.95)

```

```{r , echo=FALSE, fig.cap = "**Figure 10**. Time-series of SMU abundances (summation of all indicators stocks) relative to LRP, with all indicator stocks included."}

plotWCVI_SMUtimeseries(SMU_Esc=yy$SMU_Esc, out=yy$out, WCVI_Esc=yy$WCVIEsc)

```

