---
title: "WCVI Empirical Binomal LRPs"
author: "Carrie Holt"
date: "Aug. 6, 2021"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))  


library(tidyverse)
library(ggplot2)
library(gsl)
library(TMB)
library(viridis)

```

This report contains preliminary LRPs for WCVI Chinook based on an empirical Binomial logistic regression. Only indicators stocks with average PNI > 0.5 were included in these preliminary analyses. Results and LRPs with all indicator stocks (enhanced and unenhanced) are included at the end.  Non-enhanced indicator stocks are: Artlish, Bedwell/Ursus, Cayeghle,Kaouk, Marble, Megin, Moyeha, Nahmint, Tahsis, and Tahsish. Enhanced indicators are: Burman,  Conuma, Leiner, Nitinat, Sarita,  Somass,  Zeballos,  San Juan, Tranquil, and Gold. Note, Tranquil is considered enhanced although PNI data are not available for this stock, and has been moved. Also, Gold have been moved to the enhanced category. Although a large portion of the fish in this system are not marked (PNI =0.52), most of the unmarked spawners are thought to be second generation (or descendants of) hatchery-origin fish from Robertson Creek hatchery. There is no evidence of the original natural-spawners in this system.


```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 1**. Time series of spawner abundances by inlet for indicator WCVI Chinook stocks with average PNI >0.5. Three reference points are included: SREP from the watershed-area model (horizontal green line) with 95% prediction intervals (green shaded area), and SMSY (horizontal yellow line) and SGEN (horizontal blue line) from watershed-area SREP values and expert-derived estimates of productivity."}
source("R/helperFunctions.r")
source("R/WCVILRPs.r")
source("R/PlotSR.r")
source("R/WCVILRPs_bootstrap.R")

xx <- Get.LRP(remove.EnhStocks=TRUE)

plotWCVI_timeseries(WCVIEsc=xx$WCVIEsc, remove.EnhStocks = TRUE)


```
What is the inlet-level status?
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 2**. Annual Inlet Status, where 0 is red status and 1 is amber or green status"}

par(mfrow=c(5,1), cex=0.5,  mar = c(2, 4, 2, 1) + 0.1)
Years <- rownames(xx$WCVIEsc)

for (i in c(4,2,5,1,3)){
  plot(x=Years, y=xx$Inlet_Status[,i], pch=20, xlab="", ylab="", ylim=c(0,1), las=1 , yaxp=c(0,1,1))
  mtext(colnames(xx$Inlet_Status)[i], side=3, line=0.5, at="1952", adj=0, cex=0.8)
  if(i==5) mtext("Inlet Status = Red(0) or Amber/Green(1)", side=2, line=2, cex=1)

}


```


For each year, CU-level status is red when any one inlet within the CU is below SGEN (lower benchmark) in that year.  If all inlets within a CU are above their lower benchmark, then status is amber or green.

```{r , echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 3**. Annual CU Status, where 0 is red status and 1 is amber or green status"}

par(mfrow=c(3,1), cex=1.1,  mar = c(2, 5, 2, 1) + 0.1)
Years <- rownames(xx$WCVIEsc)

for (i in c(2,1,3)){
  plot(x=Years, y=xx$CU_Status[,i], pch=20, xlab="", ylab="", ylim=c(0,1), las=1 , yaxp=c(0,1,1))
  mtext(colnames(xx$CU_Status)[i], side=3, line=0.5, at="1952", adj=0, cex=1.2)
  if(i==1) mtext("CU Status = Red(0) or Amber/Green(1)", side=2, line=2, cex=1.4)

}

```

SMU-level status is identified from the proportion of CUs that are in the red zone. Only years when status of all three CUs are available are plotted and used in the logistic regression to derive an LRP (n=13).
```{r , echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 4**. Time-series of the proportion of CUs within the SMU that are in the red zone"}


par(mfrow=c(1,1), cex=1.1,  mar = c(2, 3, 2, 1) + 0.1)
plot(x=Years, y=xx$SMU_ppn, pch=20, xlab="", ylab="", ylim=c(0,1), las=1)
mtext("WCVI SMU", side=3, line=0.5, at="1952", adj=0, cex=1.4)

```

The LRP is derived from a logistic regression of aggregate abundances to the SMU (summation of indicator stocks with average PNI > 0.5)  versus the proportion of CUs above the red zone. A penalty is added to the estimation procedure to ensure the y-intercept of the curve is near zero. A provisional LRP at 95% of CUs above their lower benchmark is plotted, but alternatives can be considered (e.g., 2/3 of CUs). This logistic regression has a relatively poor fit to the data; in one year aggregate abundances are > LRP when only 1/3 CUs > red zone, and in 2 years aggregate abundances are < LRP but all CUs are > red zone. Diagnostic statistics and plots are in development.

```{r, echo=FALSE, fig.cap ="**Figure 5**. Logistic regression of aggregate abundances vs proportion of CUs above the red zone, with an LRP at 95% of CUs above red zone (vertical dashed line, +/- 95% CIs = vertical dotted lines). Grey band are the confidence intervals for the fitted proportions of CUs."}

plotLogistic(Data=xx$out$Logistic_Data, Preds=xx$out$Preds, LRP=xx$out$LRP, useGenMean = F, plotName="WCVI_logReg_noEnh", outDir="DataOut", p=0.95)

```

Figure 4 shows that the aggregate SMU abundances are entirely below LRP. Note, the 95% confidence interval for LRP in Figure 4 is derived from uncertaity in the statistical fit of the logisitic regression, and does not include uncertainty from the underlying lower benchmarks, which are themsleves derived from the watershed-area regression of SREP and plausible ranges of productivity (Ricker a). 

To capture uncertainties in underlying benchmarks, the LRP estimation procedure was re-iterated over multiple random samples of SREP from the watershed-area model (capturing uncertainty in SREP) and samples of productivity (Ricker a) from plausible ranges derived from expert opinion. The uncertainties from individual LRP estimates were then integrated with uncertainty derived from multiple iterations of LRPs by drawing samples from the joint distribution.
```{r , echo=FALSE, include=FALSE}

nBS <- 200 # number trials for bootstrapping
outBench <- list() 
  
for (k in 1:nBS) {
  out <- Get.LRP.bs()
  
  # Save LRPs for each bootstrap
  outLRP <- as.data.frame(out$out$LRP) 
  if(k==1) LRP.bs <- data.frame(fit=outLRP$fit, upr=outLRP$upr, lwr=outLRP$lwr)
  if(k>1) LRP.bs <- add_row(LRP.bs, outLRP)
  
  # Save benchmarks for each boostraap  
  outBench[[k]] <- out$bench
}
 
  # # Is 200 enough trials? Yes
  # running.mean <- cumsum(LRP.bs$fit) / seq_along(LRP.bs$fit) 
  # plot(running.mean)
  
# Calculate distribution of overall LRPs by integrating bootstrapped LRP 
# values with uncertainty of each LRP value from TMB
LRP.samples <- rnorm(nBS*10, LRP.bs$fit, (LRP.bs$fit - LRP.bs$lwr) / 1.96)
hist(LRP.samples)
LRP.boot <- quantile(LRP.samples, probs=c(0.05, 0.5, 0.95))
names(LRP.boot) <- c("lwr", "LRP", "upr")


nameStocks <- read.csv("DataOut/WCVI_SMSY_noEnh_wBC.csv") %>% 
  filter(Stock != "Cypre") %>% select(Stock)
nameStocks <- unique(nameStocks)

  
    
# Compile bootstrapped estimates of Sgen, SMSY, and SREP, and identify 5th and 
# 95th percentiles
SGEN.bs <- select(as.data.frame(outBench), starts_with("SGEN"))
rownames(SGEN.bs) <- nameStocks$Stock
SGEN.boot <- data.frame(SGEN= apply(SGEN.bs, 1, quantile, 0.5), 
                        lwr=apply(SGEN.bs, 1, quantile, 0.05),
                        upr=apply(SGEN.bs, 1, quantile, 0.95) )
  
SMSY.bs <- select(as.data.frame(outBench), starts_with("SMSY"))
rownames(SMSY.bs) <- nameStocks$Stock
SMSY.boot <- data.frame(SMSY= apply(SMSY.bs, 1, quantile, 0.05), 
                        lwr=apply(SMSY.bs, 1, quantile, 0.05),
                        upr=apply(SMSY.bs, 1, quantile, 0.95) )
  
SREP.bs <- select(as.data.frame(outBench), starts_with("SREP"))
rownames(SREP.bs) <- nameStocks$Stock
SREP.boot <- data.frame(SREP= apply(SREP.bs, 1, quantile, 0.5), 
                        lwr=apply(SREP.bs, 1, quantile, 0.05),
                        upr=apply(SREP.bs, 1, quantile, 0.95) )

boot <- list(LRP.boot=LRP.boot, SGEN.boot=SGEN.boot, SMSY.boot=SMSY.boot, SREP.boot=SREP.boot)


```

```{r fig.cap = "**Figure 6**. Time-series of inlet abundances (summation of indicators stocks with average PNI >0.5) relative to watershed-area based benchmarks with bootstrapped CIs."}


plotWCVI_timeseries(WCVIEsc=xx$WCVIEsc, remove.EnhStocks = TRUE, boot=boot)

```

```{r fig.cap = "**Figure 6**. Time-series of SMU abundances (summation of indicators stocks with average PNI >0.5) relative to empirical (binomial) LRP."}

LRP_boot <- data.frame(lwr=LRP.boot["lwr"], fit=LRP.boot["LRP"], upr=LRP.boot["upr"])
row.names(LRP_boot) <- "LRP"
plotWCVI_SMUtimeseries(SMU_Esc=xx$SMU_Esc, out=LRP_boot, WCVI_Esc=xx$WCVIEsc)


```


One limitation with empirical LRPs based on binomial regression is that LRPs are very sensitive to the choice of proportion of CUs near 100% (e.g., 95%, 99%, 99.9%), and cannot be estimated at 100%. alternately, empirical LRPs can be derived from Bernoulli logistic regression where the choice of the proportion is chosen a priori, and can be 100%. THe Bernoulli regression estimates the probability that the specified proportion of CUs are greater than their lower benchmarks for different aggregate abundances. The Bernoulli regression is recommended over the Binomial regression as it can be estimated for 100% of CUs> lower benchmark, however, for WCVI Chinook, it cannot be estimated as there are no years when all CUs are > lower benchmarks (Fig. 5, enhanced stocks removed).

Projection-based LRPs can be estimated based on assumed stock-recruitment parameters from watershed-area model, independent life-stage specific modeling and expert opinion. THe following logistic curve is based projections of CU-specific population dynamics over 100 years to identify aggregate abundances with at least 50% probability of all component CUs being above lower (watershed-area-based) benchmarks. In this example, the stock parameters were assumed to be known and consistent with those used to derive lower benchmarks from the watershed-area models (nTrials=100). This is a preliminary projection-based LRP value (to be updated); projection-based LRPs have not been developed for SMU with enhanced stocks. 

```{r fig.cap = "**Figure 7**. Time-series of SMU abundances (summation of indicators stocks with average PNI >0.5) relative to projection-based LRP."}
# Calculate projection-based LRP from SalmonLRP_RetroEval git repo
projLRP <- data.frame(read.csv("c:/github/SalmonLRP_RetroEval/WCVIChinookStudy/DataOut/ProjectedLRPs/ProjectedLRPsbaseER_ALLp.csv"))
projLRP <- projLRP %>% filter(ProbThresh=="0.5") %>% pull(LRP)# To Fix once completed projection-based LRPs
projLRP <- data.frame(fit=projLRP)
plotWCVI_SMUtimeseries(SMU_Esc=xx$SMU_Esc, out=projLRP, WCVI_Esc=xx$WCVIEsc)


#plotWCVI_SMUtimeseries(SMU_Esc=yy$SMU_Esc, out=yy$out$LRP, WCVI_Esc=yy$WCVIEsc)

```




## Plots with enhancement
```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 8**. Time series of spawner abundances by inlet including all indicator WCVI Chinook stocks. Three reference points are included: SREP from the watershed-area model (horizontal green line) with 95% prediction intervals (green shaded area), and SMSY (horizontal yellow line) and SGEN (horizontal blue line) estimated from SREP and expert derived productivity. The reference points in this plot differ from those in Fig. 1 because all indicator stocks were included in the watershed-area model instead of only un-enhanced indicator stocks."}
yy <- Get.LRP(remove.EnhStocks=FALSE)

plotWCVI_timeseries(WCVIEsc=yy$WCVIEsc, remove.EnhStocks = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "**Figure 9**. Annual CU Status with enhancement, where 0 is red status and 1 is amber or green status"}

par(mfrow=c(3,1), cex=1.1,  mar = c(2, 5, 2, 1) + 0.1)
Years <- rownames(yy$WCVIEsc)

for (i in c(2,1,3)){
  plot(x=Years, y=yy$CU_Status[,i], pch=20, xlab="", ylab="", ylim=c(0,1), las=1 , yaxp=c(0,1,1))
  mtext(colnames(yy$CU_Status)[i], side=3, line=0.5, at="1952", adj=0, cex=1.2)
  if(i==1) mtext("CU Status = Red(0) or Amber/Green(1)", side=2, line=2, cex=1.4)

}

```
```{r , echo=FALSE, fig.cap ="**Figure 10**. Time-series of proportion of CUs within the SMU in the red zone, with enhancement"}


par(mfrow=c(1,1), cex=1.1,  mar = c(2, 3, 2, 1) + 0.1)
plot(x=Years, y=yy$SMU_ppn, pch=20, xlab="", ylab="", ylim=c(0,1), las=1)
mtext("WCVI SMU", side=3, line=0.5, at="1952", adj=0, cex=1.4)


```

```{r , echo=FALSE, fig.cap ="**Figure 11**. Logistic regression of aggregate abundances vs proportion of CUs above the red zone, with an LRP at 95% of CUs above red zone (vertical dashed line, +/- 95% CIs = vertical dotted lines), including enhanced stocks. Grey band are the confidence intervals for the fitted proportions of CUs."}

plotLogistic(Data=yy$out$Logistic_Data, Preds=yy$out$Preds, LRP=yy$out$LRP, useGenMean = F, plotName="WCVI_logReg_noEnh", outDir="DataOut", p=0.95)

```

```{r , echo=FALSE, include=FALSE}

nBS <- 200 # number trials for bootstrapping
outBench <- list() 
  
for (k in 1:nBS) {
  out <- Get.LRP.bs(remove.EnhStocks=FALSE)
  
  # Save LRPs for each bootstrap
  outLRP <- as.data.frame(out$out$LRP) 
  if(k==1) LRP.bs <- data.frame(fit=outLRP$fit, upr=outLRP$upr, lwr=outLRP$lwr)
  if(k>1) LRP.bs <- add_row(LRP.bs, outLRP)
  
  # Save benchmarks for each boostraap  
  outBench[[k]] <- out$bench
}

  
# Calculate distribution of overall LRPs by integrating bootstrapped LRP 
# values with uncertainty of each LRP value from TMB
LRP.samples <- rnorm(nBS*10, LRP.bs$fit, (LRP.bs$fit - LRP.bs$lwr) / 1.96)
hist(LRP.samples)
LRP.boot <- quantile(LRP.samples, probs=c(0.05, 0.5, 0.95))
names(LRP.boot) <- c("lwr", "LRP", "upr")


nameStocks <- read.csv("DataOut/WCVI_SMSY_wEnh_wBC.csv") %>% 
  filter(Stock != "Cypre") %>% select(Stock)
nameStocks <- unique(nameStocks)

  
    
# Compile bootstrapped estimates of Sgen, SMSY, and SREP, and identify 5th and 
# 95th percentiles
SGEN.bs <- select(as.data.frame(outBench), starts_with("SGEN"))
rownames(SGEN.bs) <- nameStocks$Stock
SGEN.boot <- data.frame(SGEN= apply(SGEN.bs, 1, quantile, 0.5), 
                        lwr=apply(SGEN.bs, 1, quantile, 0.05),
                        upr=apply(SGEN.bs, 1, quantile, 0.95) )
  
SMSY.bs <- select(as.data.frame(outBench), starts_with("SMSY"))
rownames(SMSY.bs) <- nameStocks$Stock
SMSY.boot <- data.frame(SMSY= apply(SMSY.bs, 1, quantile, 0.5), 
                        lwr=apply(SMSY.bs, 1, quantile, 0.05),
                        upr=apply(SMSY.bs, 1, quantile, 0.95) )
  
SREP.bs <- select(as.data.frame(outBench), starts_with("SREP"))
rownames(SREP.bs) <- nameStocks$Stock
SREP.boot <- data.frame(SREP= apply(SREP.bs, 1, quantile, 0.5), 
                        lwr=apply(SREP.bs, 1, quantile, 0.05),
                        upr=apply(SREP.bs, 1, quantile, 0.95) )

boot <- list(LRP.boot=LRP.boot, SGEN.boot=SGEN.boot, SMSY.boot=SMSY.boot, SREP.boot=SREP.boot)

```

```{r , echo=FALSE, fig.cap ="**Figure 12**. Time-series of inlet abundances (summation of all indicators stocks including enhanced indicators) relative to watershed-area based benchmarks with bootstrapped CIs."}
plotWCVI_timeseries(WCVIEsc=yy$WCVIEsc, remove.EnhStocks = FALSE, boot=boot)
```

```{r,  echo=FALSE, fig.cap ="**Figure 13**. Time-series of SMU abundances (summation of all indicators stocks including enhanced indicators) relative to empirical (binomial) LRP."}
LRP_boot <- data.frame(lwr=LRP.boot["lwr"], fit=LRP.boot["LRP"], upr=LRP.boot["upr"])
row.names(LRP_boot) <- "LRP"
plotWCVI_SMUtimeseries(SMU_Esc=yy$SMU_Esc, out=LRP_boot, WCVI_Esc=yy$WCVIEsc)
```

